---
title: "HERD 2018 Notebook"
author: "Jason P. Casey"
date: 2018-11-20
output: html_notebook
---

# Setup

Data setup:




THIS _nsf_ipeds_bridge_ SCRIPT MUST BE RUN PRIOR TO RUNNING THIS SCRIPT





```{r setup}
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

system.time({
  herd <- read_db("ched-staging", "SELECT * FROM dbo.raw_herd")
})
```

# Inspect Contents of HERD Table

```{r}
herd %>%
  count(questionnaire_no, question, row, column)
```

# Create Data Tables

## Institution Level Data Table

Institutional totals for things like Medical School Expenditures, Total R&D, etc.

```{r}
fix_null <- function(x) {
  ifelse(is.na(x), 0, x)
}

inst_data <- herd %>%
  filter(questionnaire_no %in% c("01.a", "01.b", "01.c", "01.d", "01.e", "01.f", "01.g",
                                 "04")) %>%
  rename(unitid = unitid_xwalk) %>%
  select(inst_id:inst_zip, questionnaire_no, data) %>%
  mutate(data = data * 1000) %>%
  spread(key = "questionnaire_no", value = "data") %>%
  mutate(date_key = (year_id * 10000) + 0630,
         federal_government = fix_null(`01.a`),
         state_and_local_government = fix_null(`01.b`),
         business = fix_null(`01.c`),
         nonprofit_organizations = fix_null(`01.d`),
         institutional_funds = fix_null(`01.e`),
         other_sources = fix_null(`01.f`),
         total_rd_expenses = fix_null(`01.g`),
         medical_school_expenses = fix_null(`04`),
         non_medical_rd = total_rd_expenses - medical_school_expenses) %>%
  select(inst_id, ncses_inst_id, unitid, date_key, hbcu_flag:inst_zip, federal_government:non_medical_rd)

inst_data
```

## Detail Data Table

Data by academic field and funding agency.

```{r}
herd_details <- herd %>%
  mutate(data = ifelse(is.na(data), 0, data * 1000),
         date_key = (year_id * 10000) + 0630) %>%
  filter(column %in% c("DOD",
                       "DOE",
                       "HHS",
                       "NASA",
                       "NSF",
                       "USDA",
                       "Other agencies",
                       "State and local government",
                       "Business",
                       "Institution funds",
                       "Nonprofit organziations")) %>%
  separate(questionnaire_no, c("rownum","academic_field_key"), 2) %>%
  filter(as.integer(rownum) %in% c(9, 11),
         academic_field_key %in% c('A','B01','B02','B03','B04','B05','B06','B07','B08',
                                  'B09','C01','C02','C03','C04','D01','D02','D03','D04',
                                  'D05','E','F01','F02','F03','F04','F05','G','H01',
                                  'H02','H03','H04','H05','I','J01','J02','J03','J04',
                                  'J05','J06','J07','J08')) %>%
  mutate(funding_type = recode(as.integer(rownum),
                `9`="Federal",
                `11`="Non-federal"),
         agency_key = recode(column,
                             "DOD" = "DOD",
                             "DOE" = "DOE",
                             "HHS" = "HHS",
                             "NASA" = "NAS",
                             "NSF" = "NSF",
                             "USDA" = "USD",
                             "Other agencies" = "OTH",
                             "State and local government" = "SLG",
                             "Business" = "BUS",
                             "Institution funds" = "INS",
                             "Nonprofit organziations" = "NPO",
                             .default = "Unknown")) %>%
  rename(unitid = unitid_xwalk) %>%
  select(inst_id, ncses_inst_id, unitid, date_key, funding_type, agency_key, academic_field_key, expenditure = data)  %>%
  group_by(inst_id, ncses_inst_id, unitid, date_key, funding_type, agency_key, academic_field_key) %>%
  summarise(expenditure = sum(expenditure, na.rm = TRUE)) %>%
  ungroup()

herd_details
```

# Write Records to Database

## Institution Data

```{r}
inst_data %>%
  write_db("ched-staging", "nsf_herd_institution_data")
```

## Detail Data

```{r}
herd_details %>%
  write_db("ched-staging", "nsf_herd_detail_data")
```



