---
title: "HERD Bridge Notebook"
author: "Jason P. Casey"
date: 2018-12-11
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2010
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup



THIS SCRIPT MUST BE RUN PRIOR TO RUNNING THE HERD SCRIPT



Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

The code creates a table named _herd_ that contains raw responses for the sequence of years selected.

```{r setup}
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

if (params$read_raw == "Y") {
  xwalk <- read_db("ched-staging", "SELECT inst_id, unitid_xwalk FROM dbo.nsf_ipeds_bridge")
  
  herd_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%  
    mutate(path = str_c("https://www.nsf.gov/statistics/herd/data/csv/herd_", year, ".csv"),
           data = map(path, net_load)
           )
  
  herd <- herd_meta %>%
    select(data) %>%
    unnest() %>%
    mutate(inst_id = as.integer(inst_id)) %>%
    inner_join(xwalk, by = "inst_id") %>%
    mutate(year_id = as.integer(year),
           unitid = as.integer(ipeds_unitid),
           unitid = ifelse(is.na(unitid), as.integer(RollupId), as.integer(unitid)),
           unitid_xwalk = unitid,
           hbcu_flag = as.integer(hbcu_flag),
           med_sch_flag = as.integer(as.logical(med_sch_flag)),
           hhe_flag = as.integer(hhe_flag),
           toi_code = as.integer(toi_code),
           hdg_code = as.integer(hdg_code),
           tod_code = as.integer(toc_code)) %>%
    select(inst_id, ncses_inst_id, unitid_xwalk, year_id, hbcu_flag:column, tod_code, data)

  herd %>%
    write_db("ched-staging", "raw_herd")

    rm(herd_meta, herd, xwalk)
}
  
system.time({
  herd <- read_db("ched-staging", "SELECT * FROM dbo.raw_herd")
  institutions <- read_db("ched-staging", "SELECT * FROM dbo.institutions")
})

```

## Unique Institutions for Bridge Table

```{r}
unique <- herd %>%
  group_by(unitid_xwalk, inst_id, ncses_inst_id, inst_name_long, inst_state_code, med_sch_flag) %>%
  summarise(last_year_reported = max(year_id)) %>%
  ungroup()

mismatch <- unique %>%
  anti_join(institutions, by = c("unitid_xwalk" = "unitid"))

mismatch %>%
  select(unitid_xwalk, inst_id, inst_name_long, inst_state_code, med_sch_flag)

```

# Write to Database

```{r}
unique %>%
  select(unitid_xwalk:med_sch_flag, last_reported_year_id = last_year_reported) %>%
  write_db("ched-staging", "nsf_ipeds_bridge")
```

