---
title: "Load IPEDS Student Aid by Academic Year"
author: "Jason P. Casey"
date: 2018-12-04
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2010
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.


```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

make_year_string <- function(year) {
  str_c(
    str_extract(as.character(year - 1), "\\d{2}$"),
    str_extract(as.character(year), "\\d{2}$")
  )
}

read_file <- function(url, file_name) {
  recordset <- net_load_zip(url, file_name)

  drop_fields <- recordset %>%
    select(starts_with("x")) %>%
             colnames(.)

  recordset %>%
    select(-drop_fields)
}


if (params$read_raw == "Y") {
  sfa_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/SFA", make_year_string(year), ".zip"),
           data = map2(url, str_c("sfa", make_year_string(year), ".csv"), read_file))
  
  sfa <- sfa_meta %>%
    select(year, data) %>%
    unnest()
  
  sfa %>%
    write_db("ched-staging", "sfa_raw")
  
  rm(sfa_meta, sfa)
}

```

# Read Raw Data

```{r}
system.time({
  sfa <- read_db("ched-staging", "SELECT * FROM dbo.sfa_raw") %>%
    mutate(unitid = as.integer(unitid))
  })
```

# Merge Data

## Overall Aid

```{r}
div100 <- function(x) {
  x / 100
}

aid <- sfa %>%
  mutate(unitid = as.integer(unitid),
         date_key = (year * 10000) + 0630,
         total_undergraduates = as.integer(scugrad),
         full_time_freshman_cohort = as.integer(scugffn),
         total_number = as.integer(gistn2),
         total_on_campus = as.integer(giston2),
         total_off_campus_with_family = as.integer(gistwf2),
         total_off_campus_not_with_family = as.integer(gistof2),
         total_living_arrangement_unknown = as.integer(gistun2),
         cohort_paying_indistrict = as.integer(scfa11n),
         cohort_paying_instate = as.integer(scfa12n),
         cohort_paying_outstate = as.integer(scfa13n),
         cohort_paying_unknown = as.integer(scfa14n),
         awarded_grant_aid = as.integer(uagrntn),
         total_grant_aid = as.integer(uagrntt),
         awarded_federal_grants = as.integer(ofgrt_n),
         total_federal_grants = as.integer(ofgrt_t),
         awarded_state_local_grants = as.integer(sgrnt_n),
         total_state_local_grants = as.integer(sgrnt_t),
         awarded_institutional_grants = as.integer(igrnt_n),
         total_institutional_grants = as.integer(igrnt_t),
         awarded_federal_student_loans = as.integer(floan_n),
         total_federal_student_loans = as.integer(floan_t),
         awarded_other_student_loans = as.integer(oloan_n),
         total_other_student_loans = as.integer(oloan_t)) %>%
  select(unitid, date_key, total_undergraduates:total_other_student_loans)

```

## Aid by Income Band

```{r}
aid_income_band <- sfa %>%
  select(unitid, year,
         gis4n12, gis4g12, gis4t12, gis4a12,
         gis4n22, gis4g22, gis4t22, gis4a22,
         gis4n32, gis4g32, gis4t32, gis4a32,
         gis4n42, gis4g42, gis4t42, gis4a42,
         gis4n52, gis4g52, gis4t52, gis4a52,

         gis4n11, gis4g11, gis4t11, gis4a11,
         gis4n21, gis4g21, gis4t21, gis4a21,
         gis4n31, gis4g31, gis4t31, gis4a31,
         gis4n41, gis4g41, gis4t41, gis4a41,
         gis4n51, gis4g51, gis4t51, gis4a51,

         gis4n10, gis4g10, gis4t10, gis4a10,
         gis4n20, gis4g20, gis4t20, gis4a20,
         gis4n30, gis4g30, gis4t30, gis4a30,
         gis4n40, gis4g40, gis4t40, gis4a40,
         gis4n50, gis4g50, gis4t50, gis4a50,
         
         npis412, npis422, npis432, npis442, npis452,
         npis411, npis421, npis431, npis441, npis451,
         npis410, npis420, npis430, npis440, npis450) %>%
  gather(gis4n12:npis450,
         key = "variable", value = "amount") %>%
  separate(variable, c("stub", "type", "income_band_id", "index"), c(-3, -2, -1)) %>%
  select(unitid, year, stub, type, income_band_id, index, amount) %>%
  mutate(index = 2 - as.integer(index),
         date_key = ((year - index) * 10000) + 0630,
         income_band = recode(income_band_id,
                              "1" = "$0 - $30,000",
                              "2" = "$30,001 - $48,000",
                              "3" = "$48,001 - $75,000",
                              "4" = "$75,001 - $110,000",
                              "5" = "$110,001 or more"),
         variable = recode(type,
                           "a" = "average_amount",
                           "n" = "count",
                           "g" = "grant_recipients",
                           "t" = "total_amount",
                           "4" = "average_net_price"),
         amount = as.numeric(amount)) %>%
  filter(amount >= 0) %>%
  select(unitid, date_key, year, income_band, variable, amount)

aid_income_band <- aid_income_band %>%
  group_by(unitid, date_key) %>%
  summarize(year = max(year, na.rm = TRUE)) %>%
  ungroup() %>%
  select(unitid, year, date_key) %>%
  inner_join(aid_income_band) %>%
  spread(key = variable, value = amount) %>%
  select(unitid, date_key, income_band, count, grant_recipients, total_amount, average_amount, average_net_price)

aid_income_band %>%
  filter(unitid == 181464)
```

# Write to Database

## Total Aid

```{r}
aid %>%
  write_db("ched-staging", "ipeds_student_aid")
```

## Aid by Income Band

```{r}
aid_income_band %>%
  write_db("ched-staging", "ipeds_student_aid_by_income")
```
