---
title: "Load IPEDS Charges by Academic Year"
author: "Jason P. Casey"
date: 2018-12-03
output: html_notebook
params:
  year:
    label: "Report Year:"
    value: 2017
---

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
library(DBI)
library(odbc)
library(readxl)
library(lubridate)
library(kableExtra)
library(tidyverse)

# include user functions
source(file="../user-functions.R")

read_file <- function(spec1) {
  recordset <- read_csv(spec1,
                        locale = locale(encoding = "latin1")) %>%
    rename_all(tolower)
  
  drop_fields <- recordset %>%
    select(starts_with("x"), chg1tgtd, chg1fgtd, chg2tgtd, chg2fgtd, chg3tgtd, chg3fgtd) %>%
             colnames(.)

  recordset %>%
    select(-drop_fields)
}


```

# Read Data

```{r}
ay_meta <- tibble(year = 2010:2017)

ay_meta <-
  ay_meta %>%
  mutate(characteristics_file = ifelse(year > 2011,
                                       str_c("../../data/ic/ic", year, "_ay.csv"),
                                       str_c("../../data/ic/ic", year, "_ay_rv.csv")),
         data = map(characteristics_file, read_file)
         )

ay_meta

```

```{r}
ay_meta$data[[6]]
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
charges <- ay_meta %>%
  select(year, data) %>%
  unnest() %>%
  gather(chg1at0, chg1at1, chg1at2, chg1at3,
         chg2at0, chg2at1, chg2at2, chg2at3,
         chg3at0, chg3at1, chg3at2, chg3at3,
         chg1af0, chg1af1, chg1af2, chg1af3,
         chg2af0, chg2af1, chg2af2, chg2af3,
         chg3af0, chg3af1, chg3af2, chg3af3,
         chg4ay0, chg4ay1, chg4ay2, chg4ay3,
         chg5ay0, chg5ay1, chg5ay2, chg5ay3,
         chg6ay0, chg6ay1, chg6ay2, chg6ay3, key = "variable", value = "charge") %>%
  separate(variable, c("chargex", "class", "index"), c(3, 6)) %>%
  mutate(year_id = year + 1,
         index = 3 - as.integer(index),
         year = year_id - index,
         charge_type_id = recode(class,
                                 "1at" = 1,
                                 "2at" = 1,
                                 "3at" = 1,
                                 "1af" = 2,
                                 "2af" = 2,
                                 "3af" = 2,
                                 "4ay" = 3,
                                 "5ay" = 4,
                                 "6ay" = 5,
                                 .default = 0),
         charge_type = recode(class,
                                 "1at" = "Tuition",
                                 "2at" = "Tuition",
                                 "3at" = "Tuition",
                                 "1af" = "Fees",
                                 "2af" = "Fees",
                                 "3af" = "Fees",
                                 "4ay" = "Books and supplies",
                                 "5ay" = "On-campus room and board",
                                 "6ay" = "On-campus other expenses",
                                 .default = "Other"),
         student_residency = recode(class,
                                    "2at" = "In-state",
                                    "3at" = "Out-of-state",
                                    "2af" = "In-state",
                                    "3af" = "Out-of-state",
                                    .default = "In-district"),
         charge = as.numeric(charge)) %>%
  select(unitid, year_id, year, charge_type_id, charge_type, student_residency, charge)

instate <- charges %>%
  filter(charge_type_id > 2) %>%
  mutate(student_residency = "In-state")

outstate <- charges %>%
  filter(charge_type_id > 2) %>%
  mutate(student_residency = "Out-of-state")
  
charges <- charges %>%
  bind_rows(., instate, outstate)

charges <- charges %>%
  group_by(unitid, year) %>%
  summarize(year_id = max(year_id)) %>%
  ungroup() %>%
  semi_join(charges, .) %>%
  mutate(year_id = year) %>%
  select(unitid, year_id, student_residency, charge_type, charge)

charges
```

## Average Rates Table

This table contains the average tuition and fee amounts paid by typical undergraduate and graduate students.

```{r}
avg_ug <- ay_meta %>%
  select(year, data) %>%
  unnest() %>%
  gather(tuition1, tuition2, tuition3, tuition5, tuition6, tuition7,
         fee1, fee2, fee3, fee5, fee6, fee7,
         hrchg1, hrchg2, hrchg3, hrchg5, hrchg6, hrchg7,
         key = "variable", value = "charge") %>%
  separate(variable, c("type", "class"), -1) %>%
  mutate(year_id = year + 1,
         charge_type = recode(type,
                       "tuition" = "Average Tuition",
                       "fee" = "Required Fees",
                       "hrchg" = "Per Credit Hour Charge",
                       .default = "Unknown"),
         student_career = ifelse(as.integer(class) < 5, "Undergraduate", "Graduate"),
         student_residency = recode(class,
                                    "1" = "In-district",
                                    "2" = "In-state",
                                    "3" = "Out-of-state",
                                    "5" = "In-district",
                                    "6" = "In-state",
                                    "7" = "Out-of-state",
                                    .default = "Unknown"),
         charge = as.numeric(charge)) %>%
  select(unitid, year_id, charge_type, student_career, student_residency, charge) %>%
  spread(charge_type, charge)

avg_ug
  
```

## All Tuition and Fees by Career Level and Program

This table contains tuition and fee amounts for graduate programs, professional programs (e.g, Law), as well as undergraduate special programs.

```{r}

```


# Write to Database

## Detail Table

```{r}
system.time({
  charges %>%
    write_db("cdw", "ipeds_charges")
})
```

## Average Rates Table

```{r}
system.time({
  avg_ug %>%
    write_db("cdw", "ipeds_average_tuition")
})

```

