---
title: "Load IPEDS Graduation Rates"
author: "Jason P. Casey"
date: 2018-12-12
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2002
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(url, file_name) {
  recordset <- net_load_zip(url, file_name) %>%
    mutate(unitid = as.integer(unitid))
  
  if (c("dvgrhsm","dvgrhsw", "dvgraim", "dvgraiw", "dvgrapm", "dvgrapw", "dvgrbkm", "dvgrbkw", "dvgrwhm", "dvgrwhw") %in% colnames(recordset)) {
    recordset <- recordset %>%
      mutate(grhispm = dvgrhsm,
             grhispw = dvgrhsw,
             graianm = dvgraim,
             graianw = dvgraiw,
             grasiam = dvgrapm,
             grasiaw = dvgrapw,
             grbkaam = dvgrbkm,
             grbkaaw = dvgrbkw,
             grnhpim = NA,
             grnhpiw = NA,
             grwhitm = dvgrwhm,
             grwhitw = dvgrwhw)
    }

    if (c("grrace01", "grrace02") %in% colnames(recordset)) {
      recordset <- recordset %>%
        mutate(grnralm = grrace01,
               grnralw = grrace02,
               grunknm = grrace13,
               grunknw = grrace14)
    }
  
    if (c("grrace09", "grrace10", "grrace05", "grrace06", "grrace07", "grrace08", "grrace03", "grrace04") %in% colnames(recordset)) {
      recordset <- recordset %>%
        mutate(grhispm = grrace09,
               grhispw = grrace10,
               graianm = grrace05,
               graianw = grrace06,
               grasiam = grrace07,
               grasiaw = grrace08,
               grbkaam = grrace03,
               grbkaaw = grrace04,
               grnhpim = NA,
               grnhpiw = NA,
               grwhitm = grrace11,
               grwhitw = grrace12,
               gr2morm = NA,
               gr2morw = NA)
      }
  
  
  drop_fields <- recordset %>%
    select(starts_with("x")) %>%
             colnames(.)

  recordset %>%
    mutate(grnralm = as.integer(grnralm),
           grnralw = as.integer(grnralw),
           grunknm = as.integer(grunknm),
           grunknw = as.integer(grunknw),
           grhispm = as.integer(grhispm),
           grhispw = as.integer(grhispw),
           graianm = as.integer(graianm),
           graianw = as.integer(graianw),
           grasiam = as.integer(grasiam),
           grasiaw = as.integer(grasiaw),
           grbkaam = as.integer(grbkaam),
           grbkaaw = as.integer(grbkaaw),
           grnhpim = as.integer(grnhpim),
           grnhpiw = as.integer(grnhpiw),
           grwhitm = as.integer(grwhitm),
           grwhitw = as.integer(grwhitw),
           gr2morm = as.integer(gr2morm),
           gr2morw = as.integer(gr2morw)) %>%
    select(unitid, grtype, chrtstat, section, cohort,
           grnralm, grnralw, grunknm, grunknw, grhispm,
           grhispw, graianm, graianw, grasiam, grasiaw,
           grbkaam, grbkaaw, grnhpim, grnhpiw, grwhitm,
           grwhitw, gr2morm, gr2morw)
}

if (params$read_raw == "Y") {
  grs_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/gr", year, ".zip"),
           data = map2(url, str_c("gr", year, ".csv"), read_file))
  
  grs_raw <- grs_meta %>%
    select(year, data) %>%
    unnest()

  grs_raw %>%
    write_db("ched-staging", "grs_raw")

  rm(grs_meta, grs_raw)
}

```

# Read Data

```{r}
system.time({
  grs_raw <- read_db("ched-staging", "SELECT * FROM dbo.grs_raw")
  })
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
grad_rates <- grs_raw %>%
  filter(grtype %in% c(6, 7, 8, 13, 14, 15, 16, 17, 43, 44)) %>%
  mutate(year_id = year + 1,
         cohort_year_id = year - 5,
         field = recode(grtype,
                        `6` = "entering_cohort",
                        `7` = "exclusions",
                        `8` = "adjusted_cohort",
                        `13` = "completers_4_years",
                        `14` = "completers_5_years",
                        `15` = "completers_6_years",
                        `16` = "transfers",
                        `17` = "enrolled",
                        `43` = "enrolled",
                        `44` = "no_longer_enrolled",
                        .default = "unknown")) %>%
  select(unitid, year_id, cohort_year_id, field, grnralm:gr2morw) %>%
  gather(grnralm:gr2morw, key = "variable", value = "headcount") %>%
  spread(key = field,
         value = headcount,
         fill = 0) %>%
  separate(variable, c("survey", "ethnic_code", "sex"), sep = c(-5, -1)) %>%
  mutate(completers_5_years = completers_5_years + completers_4_years,
         completers_6_years = completers_6_years + completers_5_years,
         sex = recode(sex,
                      "m" = "Men",
                      "w" = "Women",
                      .default = "Unknown"),
         ethnicity = recode(ethnic_code,
                            "nral" = "Non-resident Alien",
                            "aian" = "American Indian or Alaska Native",
                            "asia" = "Asian",
                            "bkaa" = "Black or African American",
                            "hisp" = "Hispanic",
                            "nhpi" = "Native Hawaiian or Other Pacific Islander",
                            "whit" = "White",
                            "2mor" = "Two or More Races",
                            .default = "Unknown")) %>%
  select(-survey)

grad_rates %>%
  filter(unitid == 181464,
         year_id == 2018)
```

# Write to Database

## Detail Table

```{r}
grad_rates %>%
  write_db("ched-staging", "ipeds_grad_rates")
```

