---
title: "Load IPEDS Graduation Rates"
author: "Jason P. Casey"
date: 2018-12-12
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2016
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(url, file_name) {
  recordset <- net_load_zip(url, file_name) %>%
    mutate(unitid = as.integer(unitid))
  
  drop_fields <- recordset %>%
    select(starts_with("x")) %>%
             colnames(.)

  recordset %>%
    # select(-drop_fields)
    select(unitid,
           psgrtype, pgrevct, pgexclu, pgadjct,
           pgcmbac, ssrevct, ssexclu, ssadjct,
           sscmbac, nrrevct, nrexclu, nradjct,
           nrcmbac)
}

if (params$read_raw == "Y") {
  pell_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/gr", year, "_pell_ssl.zip"),
           data = map2(url, str_c("gr", year, "_pell_ssl.csv"), read_file))
  
  pell_raw <- pell_meta %>%
    select(year, data) %>%
    unnest()

  pell_raw %>%
    write_db("ched-staging", "pell_raw")

  rm(pell_meta, pell_raw)
}

```

# Read Data

```{r}
system.time({
  pell_raw <- read_db("ched-staging", "SELECT * FROM dbo.pell_raw")
  })
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
pell_grad_rates <- pell_raw %>%
  filter(psgrtype %in% c(2, 3)) %>%
  mutate(date_key = (year * 10000) + 0531,
         cohort_date_key = ((year - 6) * 10000) + 1015,
         cohort_type = recode(psgrtype,
                              `2` = "Bachelors Degree-seeking",
                              `3` = "Other Degree/Certificate-seeking",
                              .default = "Unknown")) %>%
  select(unitid, date_key, cohort_date_key, cohort_type, pgrevct:nrcmbac) %>%
  gather(pgrevct:nrcmbac, key = "variable", value = "headcount") %>%
  separate(variable, c("aid_type", "field"), sep = 2) %>%
  mutate(aid_type = recode(aid_type,
                           "pg" = "Pell Grant Recipients",
                           "ss" = "Subsidized Loan Recipients",
                           "nr" = "Did Not Receive Pell Grant or Subsidized Loan",
                           .default = "Unknown"),
         field = recode(field,
                        "revct" = "revised_cohort",
                        "exclu" = "exclusions",
                        "adjct" = "adjusted_cohort",
                        "cmbac" = "completers_6_years",
                        .default = "unknown")) %>%
  spread(key = field,
         value = headcount,
         fill = 0) %>%
  select(unitid, date_key, cohort_date_key, cohort_type, aid_type, adjusted_cohort, completers_6_years)


pell_grad_rates %>%
  filter(unitid == 181464,
         date_key == 20170531)
```

# Write to Database

## Detail Table

```{r}
pell_grad_rates %>%
  write_db("ched-staging", "ipeds_pell_grad_rates")
```

