---
title: "Load IPEDS Retention"
author: "Jason P. Casey"
date: 2018-12-12
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "N"
  start_year:
    label: "Start Year"
    value: 2005
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(url, file_name) {
  recordset <- net_load_zip(url, file_name)
  
  drop_fields <- recordset %>%
    select(starts_with("x")) %>%
             colnames(.)

  recordset %>%
    select(-drop_fields)
}

if (params$read_raw == "Y") {
  ay_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/ef", year, "d.zip"),
           data = map2(url, str_c("ef", year, "d.csv"), read_file))
  
  retention_raw <- ay_meta %>%
  select(year, data) %>%
  unnest()
  
  retention_raw %>%
    mutate(unitid = as.integer(unitid)) %>%
    write_db("ched-staging", "retention_raw")

  rm(ay_meta, retention_raw)
}

```

# Read Data

```{r}
system.time({
  retention_raw <- read_db("ched-staging", "SELECT * FROM dbo.retention_raw") %>%
    mutate(unitid = as.integer(unitid))
  })
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
retention <- retention_raw %>%
  mutate(year_id = year + 1,
         cohort_year_id = year) %>%
  select(unitid, year_id, cohort_year_id,
         rrftcta, ret_nmf, rrptcta, ret_nmp) %>%
  gather(rrftcta:ret_nmp, key = "variable", value = "headcount") %>%
  mutate(time_status = recode(variable,
                              "rrftcta" = "Full-time",
                              "ret_nmf" = "Full-time",
                              .default = "Part-time"),
         cohort = recode(variable,
                         "rrftcta" = "adjusted_cohort",
                         "rrptcta" = "adjusted_cohort",
                         .default = "enrolled")) %>%
  spread(key = cohort,
         value = headcount,
         fill=0) %>%
  group_by(unitid, year_id, cohort_year_id, time_status) %>%
  summarise(adjusted_cohort = sum(adjusted_cohort),
            enrolled = sum(enrolled)) %>%
  ungroup() %>%
  filter(adjusted_cohort > 0)

retention %>%
  filter(unitid == 181464,
         year_id == 2017) %>%
  mutate(retention = enrolled / adjusted_cohort)
```


# Write to Database

## Detail Table

```{r}
retention %>%
  write_db("ched-staging", "ipeds_retention")
```

