---
title: "Load IPEDS Charges by Academic Year"
author: "Jason P. Casey"
date: 2018-12-03
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "N"
  start_year:
    label: "Start Year"
    value: 2012
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(url, file) {
  df <- net_load_zip(url, file)
  
  df %>%
    select(unitid, occupcat, staffcat, ftpt,
           hrnralm, hrnralw, hrunknm, hrunknw, hrhispm, hrhispw,
           hraianm, hraianw, hrasiam, hrasiaw, hrbkaam, hrbkaaw,
           hrnhpim, hrnhpiw, hrwhitm, hrwhitw, hr2morm, hr2morw)

}

# https://nces.ed.gov/ipeds/datacenter/data/EAP2017.zip

if (params$read_raw == "Y") {
  staff_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
  mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/s", year, "_oc.zip"),
         file_name = str_c("s", year, "_oc.csv"),
         data = map2(url,
                     file_name,
                     read_file))

  staff_raw <- staff_meta %>%
    select(year, data) %>%
    unnest()
  
  staff_raw %>%
    write_db("ched-staging", "staff_raw")

  rm(staff_meta, staff_raw)
}

```

# Read Data

```{r}
system.time({
  staff_raw <- read_db("ched-staging", "SELECT * FROM dbo.staff_raw")
  })
```

# Build Tables

## Fall Staff

```{r}
fall_staff <- staff_raw %>%
  filter(ftpt > 1,
         staffcat %in% c(2210, 2220, 2230, 2261, 2262, 2263, 2264, 2270,
                         2300, 2310, 2320, 2330, 2340, 2350, 2360, 2370, 2380, 2390,
                         3210, 3220, 3230, 3261, 3262, 3263, 3264, 3270, 
                         3300, 3310, 3320, 3330, 3340, 3350, 3360, 3370, 3380, 3390,
                         4410, 4420, 4430, 4440, 4450, 4460, 4470, 4480, 4490)) %>%
  gather(hrnralm:hr2morw, key = "variable", value = "headcount") %>%
  separate(variable, c("survey", "demographic_key"), sep = c(2)) %>%
  mutate(unitid = as.integer(unitid),
         occupcat = ifelse(occupcat == 210,
                           211,
                           as.integer(occupcat)),
         time_status_id = recode(ftpt,
                              `2` = "FT",
                              .default = "PT"),
         time_status = recode(ftpt,
                              `2` = "Full-time",
                              .default = "Part-time"),
         date_key = (year * 10000) + 1101,
         employee_key = str_c(occupcat, time_status_id),
         staff_raw = recode(occupcat,
                                        `211` = "Instructional staff, primarily instruction",
                                        `212` = "Instructional staff, primarily instruction, exclusively credit",
                                        `213` = "Instructional staff, primarily instruction, exclusively not-for-credit",
                                        `214` = "Instructional staff, primarily instruction, combined credit/not-for-credit",
                                        `215` = "Instructional staff - Instruction/research/public service",
                                        `220` = "Research",
                                        `230` = "Public service",
                                        `250` = "Librarians, Curators, Archivists and other teaching Instructional support",
                                        `260` = "Librarians, Curators, and Archivists",
                                        `261` = "Archivists, Curators, and Museum Technicians",
                                        `262` = "Librarians",
                                        `263` = "Library Technicians",
                                        `264` = "Detailed library occupations not available for non-degree-granting institutions",
                                        `270` = "Other teaching and Instructional Support",
                                        `300` = "Management",
                                        `310` = "Business and Financial Operations",
                                        `320` = "Computer, Engineering, and Science",
                                        `330` = "Community Service, Legal, Arts, and Media",
                                        `340` = "Healthcare Practioners and Technical",
                                        `350` = "Service Occupations",
                                        `360` = "Sales and Related Occupations",
                                        `370` = "Office and Administrative Support",
                                        `380` = "Natural Resources, Construction, and Maintenance",
                                        `390` = "Production, Transportation, and Material Moving",
                                        `400` = "Graduate Assistants",
                                        `410` = "Graduate Assistants Teaching",
                                        `420` = "Graduate Assistants Research",
                                        `430` = "Graduate Assistants Management Occupations",
                                        `440` = "Graduate Assistants Business and Financial Operations Occupations",
                                        `450` = "Graduate Assistants Computer, Engineering, and Science Occupations",
                                        `460` = "Graduate Assistants Community Service, Legal, Arts, and Media",
                                        `470` = "Graduate Assistants Library and Instructional support",
                                        `480` = "Graduate Assistants Healthcare Practioners and Technical",
                                        .default = "Unknown")) %>%
  select(unitid, date_key, employee_key, occupcat, staff_raw,
         time_status, demographic_key, headcount)

fall_staff
```

## Employee Dimension Table

```{r}
staff_dim <- fall_staff %>%
  count(employee_key, staff_raw, time_status) %>%
  select(-n)

staff_dim
```

# Write Data

## Fall Staff

```{r}
fall_staff %>%
  select(unitid, date_key, employee_key, demographic_key, headcount) %>%
  write_db("ched-staging", "ipeds_fall_staff")
```

## Employee Dimension

```{r}
staff_dim %>%
  write_db("ched-staging", "staff_codes")
```

