---
title: "Load IPEDS Degree Completions Counts"
author: "Jason P. Casey"
date: 2019-01-24
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "N"
  start_year:
    label: "Start Year"
    value: 2007
  end_year:
    label: "Ending Year"
    value: 2009
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(year, file_name) {
  url <- str_c("https://nces.ed.gov/ipeds/datacenter/data/c", year, "_a.zip")
  
  recordset <- net_load_zip(url, file_name) %>%
    mutate(unitid = as.integer(unitid))
  
  if (year < 2008) {
    recordset <- recordset %>%
      mutate(cnralm = as.integer(crace01),
             cnralw = as.integer(crace02),
             cunknm = as.integer(crace13),
             cunknw = as.integer(crace14),
             chispm = as.integer(crace09),
             chispw = as.integer(crace10),
             caianm = as.integer(crace05),
             caianw = as.integer(crace06),
             casiam = as.integer(crace07),
             casiaw = as.integer(crace08),
             cbkaam = as.integer(crace03),
             cbkaaw = as.integer(crace04),
             cnhpim = 0,
             cnhpiw = 0,
             cwhitm = as.integer(crace11),
             cwhitw = as.integer(crace12),
             c2morm = 0,
             c2morw = 0)
      
  }
  
    if (year == 2008) {
      recordset <- recordset %>%
        mutate(cnralm = as.integer(cnralm),
               cnralw = as.integer(cnralw),
               cunknm = as.integer(cunknm),
               cunknw = as.integer(cunknw),
               chispm = as.integer(dvchsm),
               chispw = as.integer(dvchsw),
               caianm = as.integer(dvcaim),
               caianw = as.integer(dvcaiw),
               casiam = as.integer(dvcapm),
               casiaw = as.integer(dvcapw),
               cbkaam = as.integer(dvcbkm),
               cbkaaw = as.integer(dvcbkw),
               cnhpim = 0,
               cnhpiw = 0,
               cwhitm = as.integer(dvcwhm),
               cwhitw = as.integer(dvcwhw))
    }
  
    if (year == 2009) {
      recordset <- recordset %>%
        mutate(cnralm = as.integer(cnralm),
               cnralw = as.integer(cnralw),
               cunknm = as.integer(cunknm),
               cunknw = as.integer(cunknw),
               chispm = as.integer(dvchsm),
               chispw = as.integer(dvchsw),
               caianm = as.integer(dvcaim),
               caianw = as.integer(dvcaiw),
               casiam = as.integer(dvcapm),
               casiaw = as.integer(dvcapw),
               cbkaam = as.integer(dvcbkm),
               cbkaaw = as.integer(dvcbkw),
               cnhpim = 0,
               cnhpiw = 0,
               cwhitm = as.integer(dvcwhm),
               cwhitw = as.integer(dvcwhw))
    }
  
    if (year > 2009) {
      recordset <- recordset %>%
        mutate(cnralm = as.integer(cnralm),
               cnralw = as.integer(cnralw),
               cunknm = as.integer(cunknm),
               cunknw = as.integer(cunknw),
               chispm = as.integer(chispm),
               chispw = as.integer(chispw),
               caianm = as.integer(caianm),
               caianw = as.integer(caianw),
               casiam = as.integer(casiam),
               casiaw = as.integer(casiaw),
               cbkaam = as.integer(cbkaam),
               cbkaaw = as.integer(cbkaaw),
               cnhpim = as.integer(cnhpim),
               cnhpiw = as.integer(cnhpiw),
               cwhitm = as.integer(cwhitm),
               cwhitw = as.integer(cwhitw),
               c2morm = as.integer(c2morm),
               c2morw = as.integer(c2morw))
    }

  recordset %>%
    select(unitid, cipcode, majornum, awlevel,
           cnralm, cnralw, cunknm, cunknw, chispm,
           chispw, caianm, caianw, casiam, casiaw,
           cbkaam, cbkaaw, cnhpim, cnhpiw, cwhitm,
           cwhitw, c2morm, c2morw)
}

if (params$read_raw == "Y") {
  c_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/c", year, "_a.zip"),
           file_name = str_c("c", year, "_a.csv"),
           data = map2(year, file_name, read_file))
  
  c_raw <- c_meta %>%
    select(year, data) %>%
    unnest()

  c_raw %>%
    write_db("ched-staging", "c_raw")

  rm(c_meta, c_raw)
}

```

# Read Data

```{r}
system.time({
  c_raw <- read_db("ched-staging", "SELECT * FROM dbo.c_raw")
  })
```

# Merge Data Into Needed Tables

## Detail Table

3 'Associate^s degree' 
5 'Bachelor^s degree' 
7 'Master^s degree' 
17 'Doctor^s degree - research/scholarship' 
18 'Doctor^s degree - professional practice' 
19 'Doctor^s degree - other' 
1 'Award of less than 1 academic year' 
2 'Award of at least 1 but less than 2 academic years' 
4 'Award of at least 2 but less than 4 academic years' 
6 'Postbaccalaureate certificate' 
8 'Post-master^s certificate'.

```{r}
degree_completions <- c_raw %>%
  mutate(date_key = (year * 10000) + 1015,
         degree_key = str_c(awlevel, majornum),
         award_level = recode(awlevel,
                              `3` = "Associate's degree",
                              `5` = "Bachelor's degree",
                              `7` = "Master's degree",
                              `9` = "Doctor's degree",
                              `10` = "First-professional degree",
                              `17` = "Doctor's degree - research/scholarship",
                              `18` = "Doctor's degree - professional practice",
                              `19` = "Doctor's degree - other",
                              `1` = "Award of less than 1 academic year",
                              `2` = "Award of at least 1 but less than 2 academic years",
                              `4` = "Award of at least 2 but less than 4 academic years",
                              `6` = "Postbaccalaureate certificate",
                              `8` = "Post-master's certificate",
                              .default = "Unknown"),
         major_number = recode(majornum,
                               `1` = "First Major",
                               `2` = "Second Major",
                               .default = "Unknown")) %>%
  select(unitid, date_key, cipcode, degree_key, award_level, major_number, cnralm:c2morw) %>%
  gather(cnralm:c2morw, key = "variable", value = "award") %>%
  separate(variable, c("survey", "demographic_key"), sep = c(1)) %>%
  select(-survey)

degree_completions %>%
  filter(unitid == 181464,
         date_key == 20171015)
```

# Degree Dimension Table

```{r}
degree_dim <- degree_completions %>%
  count(degree_key, award_level, major_number) %>%
  select(-n)

degree_dim
```

## Check for Dupes

```{r}
degree_completions %>%
  count(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  filter(n > 1)
```


## Collapsed

```{r}
degree_completions <- degree_completions %>%
  group_by(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  summarize(headcount = sum(headcount, na.rm = TRUE))
```

# Write to Database

## Detail Table

```{r}
degree_completions %>%
  write_db("ched-staging", "ipeds_degree_completions")
```

