---
title: "Load IPEDS Admissions"
author: "Jason P. Casey"
date: 2018-12-03
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2007
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

fix_column <- function(x) {
  x <- ifelse(x < 0, 0, x)
  x <- ifelse(is.na(x), 0, x)
}

read_file <- function(url, file_name) {
  recordset <- net_load_zip(url, file_name)
  
  #  admcon1, admcon2, admcon3, admcon4, admcon5, admcon6, admcon7, admcon8, admcon9,
  
  recordset %>%
    mutate(satwr25 = ifelse("satwr25" %in% colnames(.), satwr25, NA),
           satwr75 = ifelse("satwr75" %in% colnames(.), satwr75, NA),
           actwr25 = ifelse("actwr25" %in% colnames(.), actwr25, NA),
           actwr75 = ifelse("actwr75" %in% colnames(.), actwr75, NA)) %>%
    select(unitid,
           applcn, applcnm, applcnw,
           admssn, admssnm, admssnw,
           enrlt, enrlm, enrlw, enrlft,
           enrlftm, enrlftw, enrlpt,
           enrlptm, enrlptw, satnum,
           satpct, actnum, actpct, satvr25,
           satvr75, satmt25, satmt75, satwr25,
           satwr75, actcm25, actcm75, acten25,
           acten75, actmt25, actmt75, actwr25,
           actwr75) %>%
    map(~ as.integer(.)) %>%
    as_tibble()
}

if (params$read_raw == "Y") {
  adm_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = ifelse(year > 2013,
                        str_c("https://nces.ed.gov/ipeds/datacenter/data/adm", year, ".zip"),
                        str_c("https://nces.ed.gov/ipeds/datacenter/data/ic", year, ".zip")),
           data = map2(url, ifelse(year > 2013, 
                                   str_c("adm", year, ".csv"),
                                   str_c("ic", year, ".csv")),
                       read_file))
  
  adm_raw <- adm_meta %>%
    mutate(year_id = year + 1) %>%
    select(year_id, data) %>%
    unnest()
  
  adm_raw %>%
    write_db("ched-staging", "adm_raw")

  rm(adm_meta, adm_raw)
}

```

# Read Data

```{r}
system.time({
  adm_raw <- read_db("ched-staging", "SELECT * FROM dbo.adm_raw")
  })
```

# Make Data Tables

## Admissions Table



```{r}
adm <- adm_raw %>%
  select(unitid, year_id,
         applcn, applcnm, applcnw,
         admssn, admssnm, admssnw,
         enrlt, enrlm, enrlw, enrlft,
         enrlftm, enrlftw, enrlpt,
         enrlptm, enrlptw) %>%
  map(~ fix_column(.)) %>%
  as_tibble(.) %>%
  mutate(applcn = ifelse(applcn==0, applcnm + applcnw, applcn),
         admssn = ifelse(admssn==0, admssnm + admssnw, admssn),
         enrlm = ifelse(enrlm==0, enrlftm + enrlptm, enrlm),
         enrlw = ifelse(enrlw==0, enrlftw + enrlptw, enrlw),
         enrlt = ifelse(enrlt==0, enrlm + enrlw, enrlt),
         enrlft = ifelse(enrlft==0, enrlftm + enrlftw, enrlft),
         enrlpt = ifelse(enrlpt==0, enrlptm + enrlptw, enrlpt),
         applcnu = applcn - (applcnm + applcnw),
         admssnu = admssn - (admssnm + admssnw),
         enrltu = enrlt - (enrlm + enrlw),
         enrlftu = enrlft - (enrlftm + enrlftw),
         enrlptu = enrlpt - (enrlptm + enrlptw)
         )  %>%
  gather(field,
         headcount,
         applcn:enrlptu) %>%
  mutate(sex = recode(field,
                      "applcnm" = "Men",
                      "admssnm" = "Men",
                      "enrlftm" = "Men",
                      "enrlptm" = "Men",
                      "applcnw" = "Women",
                      "admssnw" = "Women",
                      "enrlftw" = "Women",
                      "enrlptw" = "Women",
                      .default = "Unknown"),
         population = recode(field,
                             "admssnu" = "admissions",
                             "admssnm" = "admissions",
                             "admssnw" = "admissions",
                             "enrlftu" = "enrolled_full_time",
                             "enrlftm" = "enrolled_full_time",
                             "enrlftw" = "enrolled_full_time",
                             "enrlptu" = "enrolled_part_time",
                             "enrlptm" = "enrolled_part_time",
                             "enrlptw" = "enrolled_part_time",
                             .default = "applications")) %>%
  filter(headcount > 0) %>%
  spread(key=population,
         value=headcount,
         fill=0) %>%
  group_by(unitid, year_id, sex) %>%
  summarise(applications = sum(applications, na.rm = TRUE),
            admissions = sum(admissions, na.rm = TRUE),
            enrolled_full_time = sum(enrolled_full_time, na.rm = TRUE),
            enrolled_part_time = sum(enrolled_part_time, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(enrolled = enrolled_full_time + enrolled_part_time) %>%
  select(unitid, year_id, sex, applications, admissions, enrolled)

  adm %>%
    filter(unitid == 181464,
           year_id == 2018)

```

## Test Scores Table

```{r}
scores <- adm_raw %>%
    select(unitid, year_id, satnum, satpct,
           actnum, actpct, satvr25, satvr75,
           satmt25, satmt75, satwr25, satwr75,
           actcm25, actcm75, acten25, acten75,
           actmt25, actmt75, actwr25, actwr75) %>%
  mutate(actpct = actpct / 100,
         satpct = satpct / 100) %>%
  gather(satnum:actwr75, key = "variable", value = "val") %>%
  mutate(test_type = recode(variable,
                         "satnum" = "SAT",
                         "satpct" = "SAT",
                         "satvr25" = "SAT",
                         "satvr75" = "SAT",
                         "satmt25" = "SAT",
                         "satmt75" = "SAT",
                         "satwr25" = "SAT",
                         "satwr75" = "SAT",
                         "actnum" = "ACT",
                         "actpct" = "ACT",
                         "actcm25" = "ACT",
                         "actcm75" = "ACT",
                         "acten25" = "ACT",
                         "acten75" = "ACT",
                         "actmt25" = "ACT",
                         "actmt75" = "ACT",
                         "actwr25" = "ACT",
                         "actwr75" = "ACT"),
         subject = recode(variable,
                            "satnum" = "Composite",
                            "satpct" = "Composite",
                            "actnum" = "Composite",
                            "actpct" = "Composite",
                            "actcm25" = "Composite",
                            "actcm75" = "Composite",
                            "satmt25" = "Math",
                            "satmt75" = "Math",
                            "actmt25" = "Math",
                            "actmt75" = "Math",
                            "satvr25" = "Verbal",
                            "satvr75" = "Verbal",
                            "satwr25" = "Writing",
                            "satwr75" = "Writing",
                            "actwr25" = "Writing",
                            "actwr75" = "Writing",
                            "acten25" = "English",
                            "acten75" = "English"),
         count = ifelse(variable %in% c("satnum",
                                        "actnum"), val, 0),
         percentage = ifelse(variable %in% c("satpct",
                                             "actpct"), val, 0),
         percentile_25 = ifelse(variable %in% c("satvr25",
                                               "satmt25",
                                               "satwr25",
                                               "actcm25",
                                               "acten25",
                                               "actmt25",
                                               "actwr25"), val, 0),
         percentile_75 = ifelse(variable %in% c("satvr75",
                                               "satmt75",
                                               "satwr75",
                                               "actcm75",
                                               "acten75",
                                               "actmt75",
                                               "actwr75"), val, 0)) %>%
  group_by(unitid, year_id, test_type, subject) %>%
  summarise(count = sum(count, na.rm = TRUE),
         percentage = sum(percentage, na.rm = TRUE),
         percentile_25 = sum(percentile_25, na.rm = TRUE),
         percentile_75 = sum(percentile_75, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(count > 0 | percentage > 0 | percentile_25 > 0 | percentile_75 > 0) %>%
  mutate(count = ifelse(count == 0, NA, count),
         percentage = ifelse(percentage == 0, NA, percentage),
         percentile_25 = ifelse(percentile_25 == 0, NA, percentile_25),
         percentile_75 = ifelse(percentile_75 == 0, NA, percentile_75))


  scores %>%
    filter(unitid == 181464,
           year_id == 2018)

```

# Write to Database

## Admissions Table

```{r}
adm %>%
  write_db("ched-staging", "ipeds_admissions")
```

## Test Scores Tables

```{r}
scores %>%
  write_db("ched-staging", "ipeds_test_scores")
```

