---
title: "Load IPEDS Graduation Rates"
author: "Jason P. Casey"
date: 2018-12-16
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "N"
  start_year:
    label: "Start Year"
    value: 2002
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

read_file <- function(year, file_name) {
  url <- str_c("https://nces.ed.gov/ipeds/datacenter/data/ef", year, "a.zip")
  
  recordset <- net_load_zip(url, file_name) %>%
    mutate(unitid = as.integer(unitid))
  
  if (year < 2008) {
    recordset <- recordset %>%
      filter(efalevel %in% c(24, 25, 31, 32, 36, 44, 45, 51, 52, 56)) %>%
      mutate(efnralm = as.integer(efrace01),
             efnralw = as.integer(efrace02),
             efunknm = as.integer(efrace13),
             efunknw = as.integer(efrace14),
             efhispm = as.integer(efrace09),
             efhispw = as.integer(efrace10),
             efaianm = as.integer(efrace05),
             efaianw = as.integer(efrace06),
             efasiam = as.integer(efrace07),
             efasiaw = as.integer(efrace08),
             efbkaam = as.integer(efrace03),
             efbkaaw = as.integer(efrace04),
             efnhpim = 0,
             efnhpiw = 0,
             efwhitm = as.integer(efrace11),
             efwhitw = as.integer(efrace12),
             ef2morm = 0,
             ef2morw = 0)
      
  }
  
    if (year == 2008) {
      recordset <- recordset %>%
        filter(efalevel %in% c(24, 25, 31, 32, 36, 44, 45, 51, 52, 56)) %>%
        mutate(efnralm = as.integer(efnralm),
               efnralw = as.integer(efnralw),
               efunknm = as.integer(efunknm),
               efunknw = as.integer(efunknw),
               efhispm = as.integer(dvefhsm),
               efhispw = as.integer(dvefhsw),
               efaianm = as.integer(dvefaim),
               efaianw = as.integer(dvefaiw),
               efasiam = as.integer(dvefapm),
               efasiaw = as.integer(dvefapw),
               efbkaam = as.integer(dvefbkm),
               efbkaaw = as.integer(dvefbkw),
               efnhpim = 0,
               efnhpiw = 0,
               efwhitm = as.integer(dvefwhm),
               efwhitw = as.integer(dvefwhw))
    }
  
    if (year == 2009) {
      recordset <- recordset %>%
        filter(efalevel %in% c(24, 31, 32, 39, 40, 44, 51, 52, 59, 60)) %>%
        mutate(efnralm = as.integer(efnralm),
               efnralw = as.integer(efnralw),
               efunknm = as.integer(efunknm),
               efunknw = as.integer(efunknw),
               efhispm = as.integer(dvefhsm),
               efhispw = as.integer(dvefhsw),
               efaianm = as.integer(dvefaim),
               efaianw = as.integer(dvefaiw),
               efasiam = as.integer(dvefapm),
               efasiaw = as.integer(dvefapw),
               efbkaam = as.integer(dvefbkm),
               efbkaaw = as.integer(dvefbkw),
               efnhpim = 0,
               efnhpiw = 0,
               efwhitm = as.integer(dvefwhm),
               efwhitw = as.integer(dvefwhw))
    }
  
    if (year > 2009) {
      recordset <- recordset %>%
        filter(efalevel %in% c(24, 31, 32, 39, 40, 44, 51, 52, 59, 60)) %>%
        mutate(efnralm = as.integer(efnralm),
               efnralw = as.integer(efnralw),
               efunknm = as.integer(efunknm),
               efunknw = as.integer(efunknw),
               efhispm = as.integer(efhispm),
               efhispw = as.integer(efhispw),
               efaianm = as.integer(efaianm),
               efaianw = as.integer(efaianw),
               efasiam = as.integer(efasiam),
               efasiaw = as.integer(efasiaw),
               efbkaam = as.integer(efbkaam),
               efbkaaw = as.integer(efbkaaw),
               efnhpim = as.integer(efnhpim),
               efnhpiw = as.integer(efnhpiw),
               efwhitm = as.integer(efwhitm),
               efwhitw = as.integer(efwhitw),
               ef2morm = as.integer(ef2morm),
               ef2morw = as.integer(ef2morw))
    }

  recordset %>%
    select(unitid, efalevel,
           efnralm, efnralw, efunknm, efunknw, efhispm,
           efhispw, efaianm, efaianw, efasiam, efasiaw,
           efbkaam, efbkaaw, efnhpim, efnhpiw, efwhitm,
           efwhitw, ef2morm, ef2morw)
}

if (params$read_raw == "Y") {
  ef_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(url = str_c("https://nces.ed.gov/ipeds/datacenter/data/ef", year, "a.zip"),
           file_name = str_c("ef", year, "a.csv"),
           data = map2(year, file_name, read_file))
  
  ef_raw <- ef_meta %>%
    select(year, data) %>%
    unnest()

  ef_raw %>%
    write_db("ched-staging", "ef_raw")

  rm(ef_meta, ef_raw)
}

```

# Read Data

```{r}
system.time({
  ef_raw <- read_db("ched-staging", "SELECT * FROM dbo.ef_raw")
  })
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
fall_enrollment <- ef_raw %>%
  mutate(date_key = (year * 10000) + 1015,
         time_status = recode(efalevel,
                              `24` = "Full-time",
                              `25` = "Full-time",
                              `31` = "Full-time",
                              `32` = "Full-time",
                              `36` = "Full-time",
                              `39` = "Full-time",
                              `40` = "Full-time",
                              .default = "Part-time"),
         career_level = recode(efalevel,
                               `32` = "Graduate",
                               `36` = "Graduate",
                               `52` = "Graduate",
                               `56` = "Graduate",
                               .default = "Undergraduate"),
         degree_seeking = recode(efalevel,
                                 `24` = "Degree-seeking",
                                 `25` = "Degree-seeking",
                                 `31` = "Non-degree-seeking",
                                 `39` = "Degree-seeking",
                                 `40` = "Degree-seeking",
                                 `44` = "Degree-seeking",
                                 `45` = "Degree-seeking",
                                 `51` = "Non-degree-seeking",
                                 `59` = "Degree-seeking",
                                 `60` = "Degree-seeking",
                                 .default = "Unknown"),
         continuation_type = recode(efalevel,
                                    `24` = "First-time",
                                    `39` = "Transfer",
                                    `40` = "Continuing",
                                    `44` = "First-time",
                                    `59` = "Transfer",
                                    `60` = "Continuing",
                                    .default = "Unknown")) %>%
  select(unitid, date_key, time_status:continuation_type, efnralm:ef2morw) %>%
  gather(efnralm:ef2morw, key = "variable", value = "headcount") %>%
  separate(variable, c("survey", "demographic_key"), sep = c(2)) %>%
  select(-survey)

fall_enrollment %>%
  filter(unitid == 181464,
         date_key == 20171015)
```

## Check for Dupes

```{r}
fall_enrollment %>%
  count(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  filter(n > 1)
```


## Collapsed

```{r}
fall_enrollment <- fall_enrollment %>%
  group_by(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  summarize(headcount = sum(headcount, na.rm = TRUE))
```

# Write to Database

## Detail Table

```{r}
fall_enrollment %>%
  write_db("ched-staging", "ipeds_fall_enrollment")
```

