---
title: "Load IPEDS Graduation Rates"
author: "Jason P. Casey"
date: 2018-12-16
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 1990
  end_year:
    label: "Ending Year"
    value: 1997
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

ef_meta <- tibble(year = 1990:1997,
                  url = c("https://nces.ed.gov/ipeds/datacenter/data/ef90_a.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef1991_a.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef1992_a.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef1993_a.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef1994_anr.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef95_anr.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef96_anr.zip",
                          "https://nces.ed.gov/ipeds/datacenter/data/ef97_anr.zip"),
                  file_name = c("ef90_a.csv",
                                "ef1991_a.csv",
                                "ef1992_a.csv",
                                "ef1993_a.csv",
                                "ef1994_anr.csv",
                                "ef95_anr.csv",
                                "ef96_anr.csv",
                                "ef97_anr.csv"))

read_file <- function(url, file_name) {

  recordset <- net_load_zip(url, file_name) %>%
    mutate(unitid = as.integer(unitid))
  
  recordset <- recordset %>%
    mutate(line = as.integer(line),
           efalevel = recode(line,
                             `1` = 24,
                             `3` = 40,
                             `4` = 40,
                             `5` = 40,
                             `6` = 40,
                             `2` = 39,
                             `7` = 31,
                             `11` = 32,
                             `12` = 32,
                             `13` = 32,
                             `9` = 36,
                             `10` = 36,
                             `15` = 44,
                             `16` = 59,
                             `17` = 60,
                             `18` = 60,
                             `19` = 60,
                             `20` = 60,
                             `21` = 51,
                             `25` = 52,
                             `26` = 52,
                             `23` = 56,
                             `24` = 56,
                             .default = 99),
           efnralm = as.integer(efrace01),
           efnralw = as.integer(efrace02),
           efunknm = as.integer(efrace13),
           efunknw = as.integer(efrace14),
           efhispm = as.integer(efrace09),
           efhispw = as.integer(efrace10),
           efaianm = as.integer(efrace05),
           efaianw = as.integer(efrace06),
           efasiam = as.integer(efrace07),
           efasiaw = as.integer(efrace08),
           efbkaam = as.integer(efrace03),
           efbkaaw = as.integer(efrace04),
           efnhpim = 0,
           efnhpiw = 0,
           efwhitm = as.integer(efrace11),
           efwhitw = as.integer(efrace12),
           ef2morm = 0,
           ef2morw = 0) %>%
    filter(efalevel %in% c(24, 25, 31, 32, 36, 39, 40, 44, 45, 51, 52, 56, 59, 60))
  
  recordset %>%
    select(unitid, efalevel,
           efnralm, efnralw, efunknm, efunknw, efhispm,
           efhispw, efaianm, efaianw, efasiam, efasiaw,
           efbkaam, efbkaaw, efnhpim, efnhpiw, efwhitm,
           efwhitw, ef2morm, ef2morw)
}


if (params$read_raw == "Y") {
  ef_meta <- ef_meta %>%
    mutate(data = map2(url, file_name, read_file))
  
  ef_raw <- ef_meta %>%
    select(year, data) %>%
    unnest()

  ef_raw %>%
    write_db("ched-staging", "ef_raw")

  rm(ef_meta, ef_raw)
}

```

# Read Data

```{r}
system.time({
  ef_raw <- read_db("ched-staging", "SELECT * FROM dbo.ef_raw")
  })
```

# Merge Data Into Needed Tables

## Detail Table

This table contains the published charges by type: tuition, fees, books and supplies, room and board (on-campus), and other expenses.

```{r}
fall_enrollment <- ef_raw %>%
  mutate(date_key = as.integer(str_c(year, "1015")),
         time_status = recode(efalevel,
                              `24` = "Full-time",
                              `25` = "Full-time",
                              `31` = "Full-time",
                              `32` = "Full-time",
                              `36` = "Full-time",
                              `39` = "Full-time",
                              `40` = "Full-time",
                              .default = "Part-time"),
         career_level = recode(efalevel,
                               `32` = "Graduate",
                               `36` = "Graduate",
                               `52` = "Graduate",
                               `56` = "Graduate",
                               .default = "Undergraduate"),
         degree_seeking = recode(efalevel,
                                 `24` = "Degree-seeking",
                                 `25` = "Degree-seeking",
                                 `31` = "Non-degree-seeking",
                                 `39` = "Degree-seeking",
                                 `40` = "Degree-seeking",
                                 `44` = "Degree-seeking",
                                 `45` = "Degree-seeking",
                                 `51` = "Non-degree-seeking",
                                 `59` = "Degree-seeking",
                                 `60` = "Degree-seeking",
                                 .default = "Unknown"),
         continuation_type = recode(efalevel,
                                    `24` = "First-time",
                                    `39` = "Transfer",
                                    `40` = "Continuing",
                                    `44` = "First-time",
                                    `59` = "Transfer",
                                    `60` = "Continuing",
                                    .default = "Unknown")) %>%
  select(unitid, date_key, time_status:continuation_type, efnralm:ef2morw) %>%
  gather(efnralm:ef2morw, key = "variable", value = "headcount") %>%
  separate(variable, c("survey", "demographic_key"), sep = c(2)) %>%
  select(-survey)

fall_enrollment %>%
  filter(unitid == 181464)
```

## Check for Dupes

```{r}
fall_enrollment %>%
  group_by(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  summarize(headcount = sum(headcount, na.rm = TRUE),
            n = n()) %>%
  filter(n > 1)
```


## Collapsed

```{r}
fall_enrollment <- fall_enrollment %>%
  group_by(unitid, date_key, time_status, career_level, degree_seeking, continuation_type, demographic_key) %>%
  summarize(headcount = sum(headcount, na.rm = TRUE))
```

# Write to Database

## Detail Table

```{r}
system.time({
  fall_enrollment %>%
    append_db("ched-staging", "ipeds_fall_enrollment")
})
```

