---
title: "Load IPEDS Institutions"
author: "Jason P. Casey"
date: 2018-11-21
output: html_notebook
params:
  read_raw:
    label: "Read Raw"
    value: "Y"
  start_year:
    label: "Start Year"
    value: 2010
  end_year:
    label: "Ending Year"
    value: 2017
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="../user-functions.R")

# Load tables
big_10 <- c(145637, 147767, 151351, 153658, 163286, 170976, 171100,
            174066, 181464, 186380, 204796, 214777, 240444, 243780)

regents <- c(181464, 153658, 153603, 126614,
             126818, 145637, 155317, 174066,
             178396, 204796, 243780)

overlap <- c(100751, 122409, 126818, 137351,
             145600, 153603, 153658, 155317,
             155399, 159391, 171128, 178396,
             181464, 196060, 207500, 209542,
             228787, 232186, 236939, 445188)

nu <- c(181464, 181428, 181747, 181394, 181215)

mo_yr <- function(date_string) {
  my_month <- str_extract(date_string, "^\\d{2}")
  my_year <- str_extract(date_string, "\\d{4}$")
  
  my_date <- lubridate::make_date(my_year, my_month, 1)
  
  answer <- str_c(my_month,
        "/",
        str_pad(days_in_month(my_date), 2, pad = "0"),
        "/",
        my_year)
  
  return(answer)
}



fix_binary <- function(x) {
  recode(x,
         `1` = 1,
         `2` = 0)
}



read_files <- function(file1, file2) {
  #https://nces.ed.gov/ipeds/datacenter/data/HD2017.zip
  url <- "https://nces.ed.gov/ipeds/datacenter/data/"
  
  spec1 <- str_c(url, file1, ".zip")
  spec2 <- str_c(url, file2, ".zip")
  
  file1 <- str_c(file1, ".csv")
  file2 <- str_c(file2, ".csv")
  
  details <- net_load_zip(spec2, file2)
  
  recordset <- net_load_zip(spec1, file1) %>%
    rename_all(tolower) %>%
    rename(institution_name = instnm,
           state_abbreviation = stabbr,
           address = addr,
           web_address = webaddr) %>%
    left_join(details, by = "unitid") %>%
    left_join(nsf_institutions, by = "unitid") %>%
    mutate(unitid = as.integer(unitid),
           regents_peer = ifelse(unitid %in% regents, 1, 0),
           is_un = ifelse(unitid %in% nu, 1, 0),
           zip = as.character(zip),
           fips = as.integer(fips),
           service_academy = ifelse(as.integer(obereg) == 0, 1, 0),
           closedat = ifelse(is.na(closedat), "", closedat),
           closedat = ifelse(str_length(closedat) < 3, "", closedat),
           closedat = ifelse(str_detect(closedat, "^\\d{2}/\\d{4}$"),
                             mo_yr(closedat),
                             closedat),
           close_date = mdy(closedat),
           institution_level = recode(as.integer(iclevel),
                                      `1` = "Four or more years",
                                      `2` = "At least 2 but less than 4 years",
                                      `3` = "Less than 2 years (below associate)",
                                      .default = "Unknown"),
           control = recode(as.integer(control),
                            `1` = "Public",
                            `2` = "Private not-for-profit",
                            `3` = "Private for-profit",
                            .default = "Unknown"),
           highest_level_offering = recode(as.integer(hloffer),
                                           `1` = "Award of less than one academic year",
                                           `2` = "At least 1, but less than 2 academic yrs",
                                           `3` = "Associate's degree",
                                           `4` = "At least 2, but less than 4 academic yrs",
                                           `5` = "Bachelor's degree",
                                           `6` = "Postbaccalaureate certificate",
                                           `7` = "Master's degree",
                                           `8` = "Post-master's certificate",
                                           `9` = "Doctor's degree",
                                           .default = "Unknown"),
           highest_degree_offering = recode(as.integer(hdegofr1),
                                            `11` = "Doctor's degree - research/scholarship and professional practice",
                                            `12` = "Doctor's degree - research/scholarship",
                                            `13` = "Doctor's degree -  professional practice",
                                            `14` = "Doctor's degree - other",
                                            `20` = "Master's degree",
                                            `30` = "Bachelor's degree",
                                            `40` = "Associate's degree",
                                            `0` = "Non-degree granting",
                                            .default = "Unknown"),
           locale = recode(as.integer(locale),
                           `11` = "City: Large",
                           `12` = "City: Midsize",
                           `13` = "City: Small",
                           `21` = "Suburb: Large",
                           `22` = "Suburb: Midsize",
                           `23` = "Suburb: Small",
                           `31` = "Town: Fringe",
                           `32` = "Town: Distant",
                           `33` = "Town: Remote",
                           `41` = "Rural: Fringe",
                           `42` = "Rural: Distant",
                           `43` = "Rural: Remote",
                           .default = "Unknown"),
           parent_id = ifelse(newid > 0, newid, unitid),
           year_closed = ifelse(as.integer(deathyr) > 0, as.integer(deathyr), NA),
           cbsa_id = as.integer(cbsa),
           cbsa_type = recode(as.integer(cbsatype),
                              `1` = "Metropolitan Statistical Area",
                              `2` = "Micropolitan Statistical Area",
                              .default = "No CBSA"),
           csa_id = as.integer(csa),
           county_fips = as.integer(countycd),
           longitude = as.numeric(longitud),
           latitude = as.numeric(latitude),
           system_member = fix_binary(as.integer(f1systyp)),
           system_name = ifelse(str_trim(as.character(f1sysnam)) %in% c("-1", "-2", "-3"),
                                NA, as.character(f1sysnam)),
           undergraduate_offering = fix_binary(as.integer(ugoffer)),
           graduate_offering = fix_binary(as.integer(groffer)),
           degree_granting = fix_binary(as.integer(deggrant)),
           open_to_public = ifelse(as.integer(openpubl)==1, 1, 0),
           landgrant = fix_binary(as.integer(landgrnt)),
           hbcu = fix_binary(as.integer(hbcu)),
           hospital = fix_binary(as.integer(hospital)),
           medical = fix_binary(as.integer(medical)),
           nsf_medical_school = ifelse(as.integer(medical_school)==1, 1, 0),
           tribal = fix_binary(as.integer(tribal)),
           postsecondary_institution = recode(as.integer(postsec),
                                              `1` = "Primarily postsecondary institution",
                                              `2` = "Not primarily postsecondary",
                                              `3` = "Not postsecondary",
                                              .default = "Unknown"),
           rotc = recode(as.integer(slo5),
                         `1` = 1,
                         `0` = 0),
           confno1 = as.integer(confno1),
           confno2 = as.integer(confno2),
           confno3 = as.integer(confno3),
           confno4 = as.integer(confno4),
           active = ifelse(as.integer(cyactive)==1, 1, 0)) %>%
    select(unitid, institution_name, address, city, zip,
           web_address, fips, county_fips, service_academy,
           institution_level, control, highest_level_offering, highest_degree_offering,
           undergraduate_offering, graduate_offering, degree_granting, 
           regents_peer, is_un, locale, parent_id, year_closed, close_date,
           cbsa_id, cbsa_type, csa_id, longitude, latitude,
           system_member, system_name, open_to_public, landgrant,
           hbcu, hospital, medical, nsf_medical_school, tribal, rotc,
           confno1, confno2, confno3, confno4
           )
}

if (params$read_raw == "Y") {
  # load dependencies
  states <- read_db("ched-staging", "SELECT * FROM states")
  institution_names <- read_excel("../../data/institution-names.xlsx")
  nsf_institutions <- read_db("ched-staging", "SELECT unitid_xwalk AS unitid, med_sch_flag AS medical_school FROM nsf_ipeds_bridge") %>%
    group_by(unitid) %>%
    summarize(medical_school = max(medical_school))

  ic_meta <- tibble(year = seq(params$start_year, params$end_year)) %>%
    mutate(data = map2(str_c("hd", year), str_c("ic", year), read_files))
  
  institutions <- ic_meta %>%
    mutate(year_id = year + 1) %>%
    select(year_id, data) %>%
    unnest()
  
  rm(ic_meta, states, institution_names, nsf_institutions)
  
  institutions %>%
    write_db("ched-staging", "ipeds_institutions")
}

```

# Read Data

```{r}
if (params$read_raw == "N") {
  system.time({
    institutions <- read_db("ched-staging", "SELECT * FROM dbo.ipeds_institutions")
    })
}

```

# Merge Data

```{r}
institutions %>%
  filter(is_un == 1)
```

