---
title: "Degree Goals"
author: "Jason P. Casey"
date: 2019-07-09
output: html_notebook
---

# Setup

Data setup:

_read_raw_ parameter determines whether to do a data pull across the web.  This is costly, so only do so if new data need to be pulled.  Two options: Y or N.

_start_year_ parameter determines the first year to be added.  Correspondingly,
_end_year_ parameter sets the ending year.  Present values are based on the universe change in 2010 and the most recent year of available data.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.width = 8,
                      fig.asp = 0.618,
                      echo = TRUE,
                      warning = FALSE)
knitr::knit_engines$set(python = reticulate::eng_python)

# include user functions
source(file="user-functions.R")

```

# Read Data

## Institution Data

```{r}
system.time({
  # c_raw <- read_db("ched-staging", "SELECT * FROM dbo.ipeds_degree_completions_fact")
  # institutions <- read_db("ched-staging", "SELECT * FROM dbo.ipeds_institution_dimension") %>%
  #   filter(control == "Public",
  #          carnegie_2015 == "Doctoral Universities: Highest Research Activity")
  # enr_raw <- read_db("ched-staging", "SELECT * FROM dbo.ipeds_fall_enrollment_fact WHERE headcount > 0")
  # demo <- read_db("ched-staging", "SELECT * FROM dbo.ipeds_demographic_dimension")
  deg <- 
    read_db("ched-staging", "SELECT * FROM dbo.ipeds_degree_dimension")
  # dates <- read_db("ched-staging", "SELECT * FROM dbo.date_dimension")
  })
```

## Dates

```{r}
system.time({
  dates <- 
    read_db("ched-staging", "SELECT * FROM dbo.date_dimension")
  })
```

## Demographics

```{r}
system.time({
  demo <- 
    read_db("ched-staging", "SELECT * FROM dbo.ipeds_demographic_dimension") %>%
    mutate(urm = ifelse(str_sub(demographic_key, 1, 4) %in% c("2mor", "aian", "bkaa", "hisp", "nhpi"), 1, 0),
         female = ifelse(sex == "Women", 1, 0))
  })
```

## Enrollment

```{r}
system.time({
  enrollment <- 
    read_db("ched-staging", "SELECT * FROM dbo.ipeds_fall_enrollment_fact WHERE unitid = 181464") %>%
    filter(career_level == "Undergraduate",
           # continuation_type %in% c("Continuing", "Transfer"),
           continuation_type == "First-time",
           degree_seeking == "Degree-seeking") %>%
    inner_join(dates) %>%
    mutate(year = year + 4) %>%
    group_by(year, demographic_key) %>%
    summarize(headcount = sum(headcount))
  })

```

# Merge Data Into Needed Tables

```{r}
system.time({
  degrees <- 
    read_db("ched-staging", "SELECT * FROM dbo.ipeds_degree_completions_fact WHERE unitid = 181464") %>%
    filter(degree_key %in% c("051", "052")) %>%
    inner_join(dates) %>%
    group_by(year, demographic_key) %>%
    summarize(bachelors = sum(awards))
  })

```

```{r}
merged_df <-
  enrollment %>%
  left_join(degrees) %>%
  inner_join(demo) %>%
  filter(headcount > 0,
         bachelors > 0) %>%
  select(year, race_ethnicity, sex, urm, female, headcount, bachelors)

merged_df
```

# Look at Relationships

## Non-white

```{r}
merged_df %>%
  filter(race_ethnicity != "White") %>%
  group_by(year, race_ethnicity) %>%
  summarize(headcount = sum(headcount),
            bachelors = sum(bachelors)) %>%
  ungroup() %>%
  ggplot(aes(x = headcount, y = bachelors, color = race_ethnicity)) +
  geom_point()
```

## White

```{r}
merged_df %>%
  filter(race_ethnicity == "White") %>%
  group_by(year, race_ethnicity) %>%
  summarize(headcount = sum(headcount),
            bachelors = sum(bachelors)) %>%
  ungroup() %>%
  ggplot(aes(x = headcount, y = bachelors, color = race_ethnicity)) +
  geom_point()
```

```{r}
merged_df %>%
  mutate(re = recode(race_ethnicity,
                     "Non-resident Alien" = "International",
                     "American Indian or Alaska Native" = "Other",
                     "Native Hawaiian or Other Pacific Islander" = "Other",
                     "Two or More Races" = "Other",
                     # "Unknown" = "Other",
                     .default = race_ethnicity)) %>%
  group_by(year, re, sex) %>%
  summarize(headcount = sum(headcount),
            bachelors = sum(bachelors)) %>%
  ungroup() %>%
  ggplot(aes(x = headcount, y = bachelors)) +
  geom_point() +
  facet_grid(re ~ sex)

```

# Run a Model

```{r}
mod <- 
  merged_df %>%
  lm(bachelors ~ headcount + urm * female, data = .)

summary(mod)
```

```{r}
plot(mod)
```

